\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3.2\baselineskip}

### 2017-03-24 paper/ms.tex
Typo :0

\citet{Ambikasaran:2015} demonstrated two key facts that allow us to use this
extended matrix formalism in practice.
First, even if the covariance function is a mixture of exponentials, the
extended matrix will still be banded with a bandwidth that scales linearly
with the number of components, $J$.
Second, \citet{Ambikasaran:2015} proved that the absolute value of the
determinant of $K_\mathrm{ext}$ is equal to the determinant of $K$.
This means we can use this extended matrix formalism to compute the
marginalized likelihood in $\mathcal{O}(N)$ operations.

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/matrix.pdf}
\caption{A pictorial representation of the sparse extended matrix
$K_\mathrm{ext}$ with $N=5$ and $J=2$.
Each colored block corresponds to a non-zero entry in the matrix as described
in the legend.
    \figurelabel{matrix}}
\end{center}
\end{figure}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/benchmark_darwin.pdf}
\caption{A benchmark showing the computational scaling of \celerite\ using the
    \texttt{vanilla} band solver as a function of the number of data points
    and the number of terms.
    \emph{(left)} The cost of computing \eq{gp-likelihood} with a covariance
    matrix given by \eq{celerite-kernel} as a function of the number of data
    points $N$.
    The different colors show the cost for different numbers of terms $J$ as
    listed in the legend.
    To guide the eye, the black line shows linear scaling in $N$.
    \emph{(right)} The same information plotted as a function of $J$ for
    different values of $N$.
    The legend shows the value of $N$ for each color and the black line shows
    quadratic scaling in $J$.
    \figurelabel{benchmark}}
\end{center}
\end{figure}

@@

For this examplewe simulate a dataset using a known GP model with a kernel
outside of the support of a \celerite\ process.
This means the true autocorrelation of the process can never be correctly
represented by the model that we're using to fit but we use this example
to demonstrate that, at least in this case, valid inferences can still be made
about the physical parameters of the model.

@@

\section{Comparisons to other methods}

>>>>

\citet{Ambikasaran:2015} demonstrated two key facts that allow us to use this
extended matrix formalism in practice.
First, even if the covariance function is a mixture of exponentials, the
extended matrix will still be banded with a bandwidth that scales linearly
with the number of components, $J$.
Second, \citet{Ambikasaran:2015} proved that the absolute value of the
determinant of $K_\mathrm{ext}$ is equal to the absolute value of the
determinant of $K$.
This means we can use this extended matrix formalism to compute the
marginalized likelihood in $\mathcal{O}(N)$ operations.

@@

\begin{figure}[tp]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/matrix.pdf}
\caption{A pictorial representation of the sparse extended matrix
$K_\mathrm{ext}$ with $N=5$ and $J=2$.
Each colored block corresponds to a non-zero entry in the matrix as described
in the legend.
    \figurelabel{matrix}}
\end{center}
\end{figure}

@@

\begin{figure}[tp]
\begin{center}
\includegraphics[width=\textwidth]{figures/benchmark_darwin.pdf}
\caption{A benchmark showing the computational scaling of \celerite\ using the
    \texttt{vanilla} band solver as a function of the number of data points
    and the number of terms.
    \emph{(left)} The cost of computing \eq{gp-likelihood} with a covariance
    matrix given by \eq{celerite-kernel} as a function of the number of data
    points $N$.
    The different colors show the cost for different numbers of terms $J$ as
    listed in the legend.
    To guide the eye, the black line shows linear scaling in $N$.
    \emph{(right)} The same information plotted as a function of $J$ for
    different values of $N$.
    The legend shows the value of $N$ for each color and the black line shows
    quadratic scaling in $J$.
    \figurelabel{benchmark}}
\end{center}
\end{figure}

@@

For this example we simulate a dataset using a known GP model with a kernel
outside of the support of a \celerite\ process.
This means the true autocorrelation of the process can never be correctly
represented by the model that we're using to fit but we use this example
to demonstrate that, at least in this case, valid inferences can still be made
about the physical parameters of the model.

@@

\newpage
\section{Comparisons to other methods}
